---
title: "P8106_HW2_SY3352"
author: "Su Yan"
date: "2026-03-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = T, message = FALSE, results='hide', warning=FALSE}
library(caret) 
library(tidymodels)
library(splines) 
library(mgcv) 
library(pdp) 
library(earth) 
library(tidyverse)
library(ggplot2)
```

## Problem a

```{r}
data <- read.csv("College.csv") |> na.omit()

set.seed(2026)
data_split <- initial_split(data, prop = 0.8)

training_data <- training(data_split)
testing_data <- testing(data_split)

fit.ss <- smooth.spline(training_data$Top10perc, training_data$Outstate)

Top10perc.grid <- seq(from = min(training_data$Top10perc), 
                      to = max(training_data$Top10perc), by = 1)
p <- ggplot(data = training_data, aes(x = Top10perc, y = Outstate)) + 
  geom_point(color = rgb(.2, .4, .2, .5))

df_values <- seq(2, 20, by = 1)
pred_list = list()

for (df_ss in df_values) {
  fit <- smooth.spline(training_data$Top10perc, training_data$Outstate, 
                       df = df_ss)
  pred <- predict(fit, x = Top10perc.grid)
  pred_list[[as.character(df_ss)]]  <- data.frame(
    Top10perc1 = Top10perc.grid, Outstate1 = pred$y, df = as.factor(df_ss))
}

pred_df_all <- bind_rows(pred_list)

p + geom_line(aes(x = Top10perc1, y = Outstate1, color = df, group = df), 
              data = pred_df_all) + theme_bw() + labs(color = "df")
```
I fitted the degree of freedom from 2 to 20, it can be seen from the plot that as the degree of freedom go larger, the fitted curve become more wiggly and overfitting. With a low degree of freedom (2 to 4), the fitted curve is very smooth and almost straight, indicating that it might be underfitting and miss the non-linear pattern. 

```{r}
fit.ss$df

pred.ss <- predict(fit.ss, x = Top10perc.grid)
pred.ss.df <- data.frame(Top10perc1 = Top10perc.grid, Outstate1 = pred.ss$y)

p + geom_line(aes(x = Top10perc1, y = Outstate1), 
              data = pred.ss.df,
              color = rgb(.8, .1, .1, 1)) + theme_bw() 
```
The selected df is `r fit.ss$df`. Generalized cross-validation is used to select the degree of freedom because it efficiently approximates leave-one-out cross-validation error and is the standard approach for smoothing spline tuning.

## Problem b
```{r}
ctrl1 <- trainControl(method = "cv", number = 10)

mars_grid <- expand.grid(degree = 1:3, nprune = 2:20)

x_train <- model.matrix(Outstate ~ ., training_data)[, -1]
y_train  <- training_data$Outstate

x_test  <- model.matrix(Outstate ~ ., testing_data)[, -1]
y_test  <- testing_data$Outstate

set.seed(2026)
mars.fit <- train(x_train, y_train,
                  method = "earth",
                  tuneGrid = mars_grid,
                  trControl = ctrl1)

ggplot(mars.fit) + theme_bw() 

mars.fit$bestTune

coef(mars.fit$finalModel) 
```
The regression function is: f^​(X)=17742.43
−0.873⋅h(Expend−15954)- 0.873 \cdot h(\text{Expend} - 15954)−0.873⋅h(Expend−15954)
−33.63⋅h(83−Grad.Rate)- 33.63 \cdot h(83 - \text{Grad.Rate})−33.63⋅h(83−Grad.Rate)
+0.000102⋅h(Apps−2097)×h(Room.Board−4150)+ 0.000102 \cdot h(\text{Apps} - 2097) \times h(\text{Room.Board} - 4150)+0.000102⋅h(Apps−2097)×h(Room.Board−4150)
−0.001144⋅h(2097−Apps)×h(Room.Board−4150)- 0.001144 \cdot h(2097 - \text{Apps}) \times h(\text{Room.Board} - 4150)−0.001144⋅h(2097−Apps)×h(Room.Board−4150)
+1.144⋅h(1400−Personal)+ 1.144 \cdot h(1400 - \text{Personal})+1.144⋅h(1400−Personal)
−3.450⋅CollegeCreighton×h(F.Undergrad−1405)- 3.450 \cdot \text{CollegeCreighton} \times h(\text{F.Undergrad} - 1405)−3.450⋅CollegeCreighton×h(F.Undergrad−1405)
+7375⋅CollegeBennington+ 7375 \cdot \text{CollegeBennington}+7375⋅CollegeBennington
−0.0954⋅h(4150−Room.Board)×h(23−perc.alumni)- 0.0954 \cdot h(4150 - \text{Room.Board}) \times h(23 - \text{perc.alumni})−0.0954⋅h(4150−Room.Board)×h(23−perc.alumni)
−4.606⋅h(2003−Accept)- 4.606 \cdot h(2003 - \text{Accept})−4.606⋅h(2003−Accept)
−1.622⋅h(Enroll−828)- 1.622 \cdot h(\text{Enroll} - 828)−1.622⋅h(Enroll−828)
+4.264⋅h(828−Enroll)+ 4.264 \cdot h(828 - \text{Enroll})+4.264⋅h(828−Enroll)
+0.000241⋅h(2306−Apps)×h(Expend−15954)+ 0.000241 \cdot h(2306 - \text{Apps}) \times h(\text{Expend} - 15954)+0.000241⋅h(2306−Apps)×h(Expend−15954)
+0.001133⋅h(2003−Accept)×h(Room.Board−3484)+ 0.001133 \cdot h(2003 - \text{Accept}) \times h(\text{Room.Board} - 3484)+0.001133⋅h(2003−Accept)×h(Room.Board−3484)

```{r}
p1 <- pdp::partial(mars.fit, pred.var = c("Grad_rate"), grid.resolution = 50) |>
  autoplot() + theme_bw() 

p1

mars_pred <- predict(mars.fit, newdata = x_test)
mars_mse  <- mean((mars.pred - y_test)^2)
mars_rmse <- sqrt(mean((mars.pred - y_test)^2))

mars_mse
mars_rmse
```
The MARS Test MSE is `r mars_mse`, the MARS Test RMSE is `r mars_rmse`.

## Problem c
```{r}

```

## Problem d

## Problem e
